---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(zeallot)
```

# zeallot

Variable assignment with zeal!

[coverage]: https://codecov.io/gh/r-lib/zeallot/branch/master/graph/badge.svg "deep fat fry"
[cran]: https://www.r-pkg.org/badges/version/zeallot "green means go!"

![alt text][coverage] ![alt text][cran]

## What's there to be excited about?

zeallot allows multiple, unpacking, or destructuring assignment in R by
providing the `%<-%` operator. With zeallot you can tighten code with explicit
variable names, unpack pieces of a lengthy list or the entirety of a small list,
destructure and assign object elements, or do it all at once.

Unpack a list of values.

```{r}
c(x, y) %<-% list(2, 3)

x
y
```

Destructure a data frame and assign its columns.

```{r}
c(cyl=, wt=) %<-% mtcars

cyl
wt
```

Unpack a nested list into nested left-hand side variables.

```{r}
c(c(x, y), z) %<-% list(list(1, 2), 3)

x
y
z
```

Destructure and partially unpack a list. "a" is assigned to `first` and the
remaining letters are grouped and assigned to one variable.

```{r}
c(first, ..rest) %<-% letters

first
unlist(rest)
```

## Getting started

Below is a simple usage example using the
[purrr](https://github.com/hadley/purrr) package and the safely function.

### Safe functions

The `purrr::safely` function returns a "safe" version of a function. The
following example is borrowed from the safely documentation. In this example a
safe version of the log function is created.

```{r}
safe_log <- purrr::safely(log)

safe_log(10)

safe_log("a")
```

A safe function always returns a list of two elements and will not throw an
error. Instead of throwing an error, the error element of the return list is set
and the value element is NULL. When called successfully the result element is
set and the error element is NULL.

Safe functions are a great way to write self-documenting code. However, dealing
with a return value that is always a list could prove tedious and may undo
efforts to write concise, readable code. Enter zeallot.

### The `%<-%` operator

With zeallot's unpacking operator `%<-%` we can unpack a safe function's return
value into two explicit variables and avoid dealing with the list return value
all together.

```{r}
c(res, err) %<-% safe_log(10)

res
err
```

The name structure of the operator is a flat or nested set of bare variable
names built with calls to `c()`. . These variables do not need to be previously
defined. On the right-hand side is a vector, list, or other R object to unpack.
`%<-%` unpacks the right-hand side, checks the number of variable names against
the number of unpacked values, and then assigns each unpacked value to a
variable. The result, instead of dealing with a list of values there are two
distinct variables, `res` and `err`.

### Further reading and examples

For more on the above example, other examples, and a thorough introduction to
zeallot check out the vignette on [unpacking
assignment](vignettes/unpacking-assignment.Rmd).

Below are links to discussions about multiple, unpacking, and destructuring
assignment in R,

* https://stackoverflow.com/questions/7519790/assign-multiple-new-variables-on-lhs-in-a-single-line-in-r
* https://stackoverflow.com/questions/1826519/how-to-assign-from-a-function-which-returns-more-than-one-value

## Installation

You can install the development version of zeallot from
[GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("r-lib/zeallot")
```

---

Thank you to Paul Teetor for inspiring me to build zeallot.

Without his encouragement nothing would have gotten off the ground.